```{r}
library(dplyr)
```

## Reading the data

```{r}
spotify_data <- read.csv("C:/Users/rhnpu/Downloads/DSRP_2023/data/Spotify_data.csv")
spotify_data
```

## 1. Filtering the observations of your dataset based on one or more variable using filter()

```{r}
filter(spotify_data, Gender == "Male", spotify_usage_period == "More than 2 years")
```

## 2. Create a smaller dataset with a subset of variables using select()

```{r}
spotify_small <- select(spotify_data, Age, spotify_usage_period, spotify_subscription_plan, premium_sub_willingness, music_recc_rating
)
spotify_small
```

## 3. Add 2 new columns to your dataset using mutate()

```{r}
mutate(spotify_small,
       less_promotion = premium_sub_willingness,
       rating = music_recc_rating)
```

## 4. Create a data table of grouped summaries on at least one numeric variable by at least one categorical variable using summarize()

```{r}
summarize(spotify_small,
          mean_rating = mean(music_recc_rating, na.rm = T),
          count = n(),
          .by = Age)
```

## 5. Reorder a data table of your choice by one or more variables using arrange()

```{r}
arrange(spotify_data, Age, spotify_usage_period)
```

## 6. Create at least one new visualization using some form of an updated dataset

```{r}
library(ggplot2)
#ggplot(data = spotify_small, aes(x = Age)) +
 # geom_histogram()
ggplot(data = spotify_small, aes(x = music_recc_rating)) +
  geom_histogram()
```
