```{r setup}
library(dplyr)

```

## We will be practicing with dplyr functions using starwars dataset:

```{r}
starwars
head(starwars)
```

## **Subset rows with filter**

filter() allows us to subset the rows of a dataset based on the values

```{r}
## Filter for rows with human Star Wars characters
filter(starwars, species == "Human")

## save to a new R object
humans <- filter(starwars, species == "Human")
humans
filter(humans, sex == "male")
```

We can use different comparison operators to filter:

-   `!` is "not"

-   `==` is "equal"

-   `|` is "or"

-   `<` is "less than"

-   `>` is "greater than"

-   `is.na()` returns TRUE if value is NA

```{r}
## return characters with a missing mass
filter(starwars, is.na(mass))
```

```{r}
## How to get all characters that are NOT human?
filter(starwars, species != "Human")
```

variable_name `%in%` vector_of_values returns TRUE if the value of variable_name is equal to one of the values in vector_of_values

```{r}
skin_color <- "grey"
skin_color %in% c("yellow", "fair", "tan") #returns false because "grey" is not in vector of colors

##How would we filter for characters with brown or blond hair?

filter(starwars, hair_color %in% c("yellow", "blond"))
```

grepl("characters_of_interest", variable_name) returns TRUE if the value of variable_name contains the characters_of_interest.

```{r}
filter(starwars, grepl("blond", hair_color))

## How would we get all characters with blue in their eyes
filter(starwars, grepl("blue", eye_color))

```

You can use `<` , `>` , `<=` , `>=` to compare numeric variables or categorical variables.

-   nominal character variables (character vectors) are ranked alphabetically

-   ordinal categorical variables (factors\_ are ranked based on the built-in order

```{r}
## starwars characters with a name starting with letters before "D"
filter(starwars, name < "D")
```

```{r}
between(5, 1, 7) # true
filter(starwars, between(mass, 50, 100))
```

### Practice!

Find all starwars characcters that

1.  Are missing a hair color
2.  Are shorter than 100cm and are not droids
3.  Were born between the years 100 and 200 (inclusive)
4.  Another of your choice.

```{r}
## Practice code

filter(starwars, is.na(hair_color))

filter(starwars, height < 100, species != "Droid")

filter(starwars, between(birth_year, 100, 200))
```

## Order rows with `arrange`

arrange() reorders rows based on the values of column(s)

```{r}
arrange(starwars, name)

arrange(starwars, species, birth_year)

arrange(starwars, species, desc(birth_year))
```

**Practice!**

1.  Arrange the starwars dataset from tallest to shortest characters
2.  Arrange the starwars dataset from shortest to tallest by species
3.  Alphabetize the Star Wars characters by their hair color then eye color

```{r}
arrange(starwars, desc(height))

arrange(starwars, species, height)

arrange(starwars, hair_color, eye_color)
```

Subset columns with select()

```{r}
select(starwars, hair_color, skin_color, eye_color)
character_colors <- select(starwars, hair_color, skin_color, eye_color)
character_colors
```

We can also select multiple columnds in a row

```{r}
# all columnds between name and eye_color (inclusive)
select(starwars, name:eye_color)
```

We can also remove columns by name

```{r}
starwars
select(starwars, -films, -vehicles)
select(starwars, -c(films: starships))
```

Helpful function

`everything()` is a shortcut for "everything else"

```{r}
select(starwars, species, everything())

## alphabetize columns
starwars_names <- sort(names(starwars))
starwars_names
select(starwars, starwars_names)
```

`starts_with("abc")` matches column names that start with "abc"

`ends_with("abc")` matches column names that end with "abc"

`contains("abc")` matches column names that contain "abc"

```{r}
select(starwars, starts_with("h"))

select(starwars, contains("or"))
```

**Practice!**

1.  Select each character's homeworld and species. What question might this subset of variables help answer?
2.  Brainstorm as many ways as possible to select `hair_color` ,`skin_color` ,`eye_color` ,`birth_year`

```{r}
select(starwars, c(homeworld, species))
select(starwars, c(hair_color, skin_color, eye_color, birth_year))
select(starwars, hair_color, skin_color, eye_color, birth_year)
select(starwars, c(hair_color:birth_year))
```

## Add new variables with `mutate`

```{r}
starwars_small <- select(starwars, height, mass, species)
starwars_small

mutate(starwars_small,
       height_m = height/100)
```

We can also create multiple variables at once and we can refer to variables we just created

```{r}
mutate(starwars_small,
       height_m = height/100,
       bmi = mass/height_m^2)

## Only keep new variables

transmute(starwars_small,
       height_m = height/100,
       bmi = mass/height_m^2)
```

Group by a variable using the `.by` parameter

```{r}
mutate(starwars_small,
       species_mass = mean(mass, na.rm = T),
       .by = species)
```

Summarize data with `summarize`

```{r}
summarize(starwars,
          mean_mass = mean(mass, na.rm = T))

```

```{r}
summarize(starwars,
          mean_mass = mean(mass, na.rm = T),
          count = n(),
          .by = species)
```
